### 2016-09-19 paper/paper101.tex
created folder for paper

\documentclass[article]{jss}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% almost as usual
\author{Earvin Balderama\\Loyola University Chicago \And 
        Taylor Trippe\\Loyola University Chicago}
\title{\pkg{hurdlr}: An \proglang{R} Package for Zero-inflated and Extremely Over-dispersed Count Data}

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Earvin Balderama, Taylor Trippe} %% comma-separated
\Plaintitle{hurdlr: An R Package for Zero-inflated and Extremely Over-dispersed Count Data} %% without formatting
\Shorttitle{\pkg{hurdlr}: An R Package for Count Data} %% a short title (if necessary)

%% an abstract and keywords
\Abstract{
  The abstract of the article.
}
\Keywords{keywords, comma-separated, not capitalized, \proglang{Java}}
\Plainkeywords{keywords, comma-separated, not capitalized, Java} %% without formatting
%% at least one keyword must be supplied

%% publication information
%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{2012-06-04}
%% \Acceptdate{2012-06-04}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Earvin Balderama\\
  Department of Mathematics and Statistics\\
  Loyola University Chicago\\
  1032 W Sheridan Rd., \\
  Chicago, IL 60660, USA\\
  E-mail: \email{ebalderama@luc.edu}\\
  URL: \url{http://math.luc.edu/~ebalderama/}\\

  Taylor Trippe\\
  Department of Mathematics and Statistics\\
  Loyola University Chicago\\
  1032 W Sheridan Rd., \\
  Chicago, IL 60660, USA\\
  E-mail: \email{ebalderama@luc.edu}\\
  URL: \url{http://math.luc.edu/~ebalderama/}
}

%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/512/507-7103
%% Fax: +43/512/507-2851

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% include your article here, just as usual
%% Note that you should use the \pkg{}, \proglang{} and \code{} commands.

%\section[About Java]{About \proglang{Java}}
%% Note: If there is markup in \(sub)section, then it has to be escape as above.

\section{Introduction}

\section{Models for count data}

\section{Double-hurdle models}

\section[Implementing hurdlr]{Implementing \pkg{hurdlr}}

\section{Discussion}

\section*{References}

\end{document}

### 2016-09-26 paper/paper101.tex
added some new latex commands in the preamble
just adding some paragraphs

@@
 \documentclass[article]{jss}
+
+\usepackage{amssymb, amsthm, amsmath}
+
+%============================================
+%============================================
+%My New Commands
+%============================================
+\newcommand{\balpha}{\mbox{\boldmath $\alpha$} }
+\newcommand{\bbeta}{\mbox{\boldmath $\beta$} }
+\newcommand{\bdelta}{\mbox{\boldmath $\delta$} }
+\newcommand{\bepsilon}{\mbox{\boldmath $\epsilon$} }
+\newcommand{\bgamma}{\mbox{\boldmath $\gamma$} }
+\newcommand{\blambda}{\mbox{\boldmath $\lambda$} }
+\newcommand{\bmu}{\mbox{\boldmath $\mu$} }
+\newcommand{\bnu}{\mbox{\boldmath $\nu$} }
+\newcommand{\bomega}{\mbox{\boldmath $\omega$} }
+\newcommand{\bphi}{\mbox{\boldmath $\phi$} }
+\newcommand{\bpsi}{\mbox{\boldmath $\psi$} }
+\newcommand{\brho}{\mbox{\boldmath $\rho$} }
+\newcommand{\bsigma}{\mbox{\boldmath $\sigma$} }
+\newcommand{\btau}{\mbox{\boldmath $\tau$} }
+\newcommand{\btheta}{\mbox{\boldmath $\theta$} }
+\newcommand{\bupsilon}{\mbox{\boldmath $\upsilon$} }
+\newcommand{\bxi}{\mbox{\boldmath $\xi$} }
+\newcommand{\bzeta}{\mbox{\boldmath $\zeta$} }
+\newcommand{\bDelta}{\mbox{\boldmath $\Delta$} }
+\newcommand{\bGamma}{\mbox{\boldmath $\Gamma$} }
+\newcommand{\bLambda}{\mbox{\boldmath $\Lambda$} }
+\newcommand{\bPhi}{\mbox{\boldmath $\Phi$} }
+\newcommand{\bSigma}{\mbox{\boldmath $\Sigma$} }
+\newcommand{\bTheta}{\mbox{\boldmath $\Theta$} }
+
+\newcommand{\bfa}{\mbox{\bf a} }
+\newcommand{\bfb}{\mbox{\bf b} }
+\newcommand{\bfc}{\mbox{\bf c} }
+\newcommand{\bfd}{\mbox{\bf d} }
+\newcommand{\bfe}{\mbox{\bf e} }
+\newcommand{\bff}{\mbox{\bf f} }
+\newcommand{\bfg}{\mbox{\bf g} }
+\newcommand{\bfh}{\mbox{\bf h} }
+\newcommand{\bfi}{\mbox{\bf i} }
+\newcommand{\bfj}{\mbox{\bf j} }
+\newcommand{\bfk}{\mbox{\bf k} }
+\newcommand{\bfl}{\mbox{\bf l} }
+\newcommand{\bfm}{\mbox{\bf m} }
+\newcommand{\bfn}{\mbox{\bf n} }
+\newcommand{\bfo}{\mbox{\bf o} }
+\newcommand{\bfp}{\mbox{\bf p} }
+\newcommand{\bfq}{\mbox{\bf q} }
+\newcommand{\bfr}{\mbox{\bf r} }
+\newcommand{\bfs}{\mbox{\bf s} }
+\newcommand{\bft}{\mbox{\bf t} }
+\newcommand{\bfu}{\mbox{\bf u} }
+\newcommand{\bfv}{\mbox{\bf v} }
+\newcommand{\bfw}{\mbox{\bf w} }
+\newcommand{\bfx}{\mbox{\bf x} }
+\newcommand{\bfy}{\mbox{\bf y} }
+\newcommand{\bfz}{\mbox{\bf z} }
+\newcommand{\bfA}{\mbox{\bf A} }
+\newcommand{\bfB}{\mbox{\bf B} }
+\newcommand{\bfC}{\mbox{\bf C} }
+\newcommand{\bfD}{\mbox{\bf D} }
+\newcommand{\bfE}{\mbox{\bf E} }
+\newcommand{\bfF}{\mbox{\bf F} }
+\newcommand{\bfG}{\mbox{\bf G} }
+\newcommand{\bfH}{\mbox{\bf H} }
+\newcommand{\bfI}{\mbox{\bf I} }
+\newcommand{\bfJ}{\mbox{\bf J} }
+\newcommand{\bfK}{\mbox{\bf K} }
+\newcommand{\bfL}{\mbox{\bf L} }
+\newcommand{\bfM}{\mbox{\bf M} }
+\newcommand{\bfN}{\mbox{\bf N} }
+\newcommand{\bfO}{\mbox{\bf O} }
+\newcommand{\bfP}{\mbox{\bf P} }
+\newcommand{\bfQ}{\mbox{\bf Q} }
+\newcommand{\bfR}{\mbox{\bf R} }
+\newcommand{\bfS}{\mbox{\bf S} }
+\newcommand{\bfT}{\mbox{\bf T} }
+\newcommand{\bfU}{\mbox{\bf U} }
+\newcommand{\bfV}{\mbox{\bf V} }
+\newcommand{\bfW}{\mbox{\bf W} }
+\newcommand{\bfX}{\mbox{\bf X} }
+\newcommand{\bfY}{\mbox{\bf Y} }
+\newcommand{\bfZ}{\mbox{\bf Z} }
+
+\newcommand{\calA}{{\cal A}}
+\newcommand{\calB}{{\cal B}}
+\newcommand{\calC}{{\cal C}}
+\newcommand{\calD}{{\cal D}}
+\newcommand{\calE}{{\cal E}}
+\newcommand{\calF}{{\cal F}}
+\newcommand{\calG}{{\cal G}}
+\newcommand{\calH}{{\cal H}}
+\newcommand{\calI}{{\cal I}}
+\newcommand{\calJ}{{\cal J}}
+\newcommand{\calK}{{\cal K}}
+\newcommand{\calL}{{\cal L}}
+\newcommand{\calM}{{\cal M}}
+\newcommand{\calN}{{\cal N}}
+\newcommand{\calO}{{\cal O}}
+\newcommand{\calP}{{\cal P}}
+\newcommand{\calQ}{{\cal Q}}
+\newcommand{\calR}{{\cal R}}
+\newcommand{\calS}{{\cal S}}
+\newcommand{\calT}{{\cal T}}
+\newcommand{\calU}{{\cal U}}
+\newcommand{\calV}{{\cal V}}
+\newcommand{\calW}{{\cal W}}
+\newcommand{\calX}{{\cal X}}
+\newcommand{\calY}{{\cal Y}}
+\newcommand{\calZ}{{\cal Z}}
+
+%\renewcommand{\Hat}{\widehat}
+%\renewcommand{\Bar}{\overline}
+%\renewcommand{\Tilde}{\widetilde}
+
+\let\originalleft\left
+\let\originalright\right
+\renewcommand{\left}{\mathopen{}\mathclose\bgroup\originalleft}
+\renewcommand{\right}{\aftergroup\egroup\originalright}
+
+\providecommand{\itbf}[1]{\textit{\textbf{#1}}} 
+\providecommand{\abs}[1]{\left\lvert#1\right\rvert} 
+\providecommand{\norm}[1]{\left\lVert#1\right\rVert}
+
+\newcommand{\cond}{\,\left\vert\vphantom{}\right.}
+\newcommand{\Cond}{\,\Big\vert\vphantom{}\Big.}
+\newcommand{\COND}{\,\Bigg\vert\vphantom{}\Bigg.}
+
+\providecommand{\paren}[1]{\left(#1\right)} 
+\providecommand{\Paren}[1]{\Big(#1\Big)}
+\providecommand{\PAREN}[1]{\bigg(#1\bigg)} 
+\providecommand{\bracket}[1]{\left[ #1 \right]} 
+\providecommand{\Bracket}[1]{\Big[ #1 \Big]} 
+\providecommand{\BRACKET}[1]{\bigg[ #1 \bigg]} 
+\providecommand{\curlybrace}[1]{\left\{ #1 \right\}} 
+\providecommand{\Curlybrace}[1]{\Big\{ #1 \Big\}} 
+\providecommand{\CURLYBRACE}[1]{\bigg\{ #1 \bigg\}}
+
+\newcommand{\Bern}{\mbox{{\sf Bern}}}
+\newcommand{\Bernoulli}{\mbox{{\sf Bernoulli}}}
+\newcommand{\Beta}{\mbox{{\sf Beta}}}
+\newcommand{\Bin}{\mbox{{\sf Bin}}}
+\newcommand{\Binomial}{\mbox{{\sf Binomial}}}
+\newcommand{\DE}{\mbox{{\sf DE}}}
+\newcommand{\Exponential}{\mbox{{\sf Exponential}}}
+\newcommand{\F}{\mbox{{\sf F}}}
+\newcommand{\Gam}{\mbox{{\sf Gamma}}}
+\newcommand{\GP}{\mbox{{\sf GP}}}
+\newcommand{\GPD}{\mbox{{\sf GPD}}}
+\newcommand{\Geom}{\mbox{{\sf Geom}}}
+\newcommand{\Geometric}{\mbox{{\sf Geometric}}}
+\newcommand{\HyperGeom}{\mbox{{\sf HyperGeom}}}
+\newcommand{\HyperGeometric}{\mbox{{\sf HyperGeometric}}}
+\newcommand{\InverseGam}{\mbox{{\sf InverseGamma}}}
+\newcommand{\InvWish}{\mbox{{\sf InvWish}}}
+\newcommand{\MVN}{\mbox{{\sf MVN}}}
+\newcommand{\NB}{\mbox{{\sf NB}}}
+\newcommand{\NegBin}{\mbox{{\sf NegBin}}}
+\newcommand{\NegativeBinomial}{\mbox{{\sf NegativeBinomial}}}
+\newcommand{\Normal}{\mbox{{\sf Normal}}}
+\newcommand{\Pois}{\mbox{{\sf Pois}}}
+\newcommand{\Poisson}{\mbox{{\sf Poisson}}}
+\newcommand{\Unif}{\mbox{{\sf Unif}}}
+\newcommand{\Uniform}{\mbox{{\sf Uniform}}}
+\newcommand{\Weibull}{\mbox{{\sf Weibull}}}
+
+%\renewcommand{\P}{{\sf P}}
+%\newcommand{\Prob}{{\sf Prob}}
+\newcommand{\median}{{\mathop{\rm median}}}
+%\newcommand{\E}{\mathsf{E}}
+%\newcommand{\V}{\mathsf{V}}
+%\newcommand{\VAR}{\mathsf{VAR}}
+%\newcommand{\COV}{\mathsf{COV}}
+
+\newcommand{\Ho}{{\calH_0}}
+\newcommand{\Hoc}{{\calH_0\colon}}
+\newcommand{\Hone}{{\calH_1}}
+\newcommand{\Honec}{{\calH_1\colon}}
+\newcommand{\Ha}{{\calH_a}}
+\newcommand{\Hac}{{\calH_a\colon}}
+\newcommand{\HA}{{\calH_A}}
+\newcommand{\HAc}{{\calH_A\colon}}
+
+\newcommand{\Ind}{\mathds{1}}
+\newcommand{\zerovect}{\mbox{\bf 0}}
+\newcommand{\onesvect}{\mbox{\bf 1}}
+\providecommand{\real}[1]{\mathbb{#1}}
+\newcommand{\Real}{\mathbb{R}}
+\newcommand{\ppd}{\mathcal{P}}
+%\DeclareMathOperator{\logit}{logit}
+%\DeclareMathOperator{\expit}{expit}
+%\DeclareMathOperator{\dint}{\displaystyle\int}
+%\DeclareMathOperator{\dsum}{\displaystyle\sum}
+
+%============================================
+%My New Commands
+%============================================
+%============================================
 
@@
 
+%============================================
 \section{Introduction}
+%============================================
 
+Count data is one of the most common and important forms of data encountered in every field of study.
+While some count data can be modeled easily with binomial or poisson likelihoods, many times the data is gathered such that a disproportionate amount of zeros appear relative to non-zeros.
+For these, a class of models called zero-inflated models were created with the reasoning that the counts came from two separate data-generating mechanisms; one that generated counts according to some discrete probability distribution (such as Poisson), and another that generated the excess amount of zeros.
+
+In this paper, we present an R package dedicated to modeling zero-inflated count data, with an emphasis on developing new methods to account for both zero-inflation and extreme over-dispersion simultaneously.
+The models in this package utilize Bayesian estimation techniques.
+Parameter estimation under a Bayesian framework is suitable because of the model's hierarchical structure.
+Model parameters were given uninformative priors and updated one-at-a-time using a Markov chain Monte Carlo (MCMC) algorithm.
+
+%============================================
 \section{Models for count data}
+%============================================
 
+%______________________________________________
+\subsection{Zero-inflated models}
+
+Zero-inflated models are widely used for count data exhibiting a significantly higher amount of zeros that would otherwise be observed given some typical count distribution.
+These are simply two-component mixture models that can separate the excess zeros in the data from the underlying (parametric) count distribution.
+A common choice for the count distribution is Poisson, but is typically replaced with negative binomial if there is over-dispersion in the data.
+%Models such as the zero-inflated poisson (ZIP) and the zero-inflated negative binomial (ZINB)
+
+%______________________________________________
+\subsection{Hurdle models}
+
+Hurdle models are similar to zero-inflated models in that they also account for the high occurrence of zeros in data.
+The difference is that for hurdle models, \textit{all} zero-count observations are assumed to have been generated by a mechanism separate of the non-zeros.
+We begin by trying to fit the negative binomial hurdle model, with the following distributional components:
+
+\begin{equation}
+\label{nbhurdle}
+  h_1(y_i | m_i,r) = 
+  \begin{cases}
+  p_i & \mbox{if } y_i=0,\\
+  (1-p_i)\cdot f(y_i | m_i,r) & \mbox{if } y_i>0,
+  \end{cases}
+\end{equation}
+where for each observation $i$, $0 \le p_i \le 1$ and $f(y_i | m_i,r)$ is the negative binomial probability function, with mean $m_i$ and overdispersion parameter $r$, truncated at 0.
+
+%============================================
 \section{Double-hurdle models}
+%============================================
 
+%For a particular species, we build a hierarchical model using the counts obtained in  areas of non-zero effort.
+Surveys are assumed to observe and record any species, so a count of 0 indicates an {\it observed} 0, not a missing observation.
+Later, we use model results to form predictions on unobserved areas.
+
+%However, the seabird count data exhibits not only zero-inflation, but also extreme over-dispersion.
+%For example, the common eider data consists of 10,736 zero-counts and 791 non-zero counts that range from 1 to 50,025.
+%For example, data on the northern gannet consists of 11,759 zero-counts and 2,248 non-zero counts that range from 1 to 1,775.
+The mean and median of the non-zero counts is 6.06 and 2, respectively, thus forming an extremely right-skewed distribution.
+% and with an average of 7 and a .
+
+In some cases, few but very large count values causes the data to be so extremely overdispersed that a simple negative binomial hurdle model proves to be inadequate.
+
+Properly fitting such an extremely right-skewed distribution is the goal of the double-hurdle model.
+In this model, a second hurdle is created to represent the transition from a distribution of ``typical" counts to a distribution of larger counts.
+Just as in the hurdle model, we assume separate data-generating processes for each of the components.
+The first hurdle represents the divide between the zero-counts and the ``typical" non-zero counts.
+Larger counts can be thought of as coming from a heavy-tailed distribution $g$ if the second hurdle is crossed.
+The double-hurdle model has the following form for observation $i$:
+%%not accounting for detection
+\begin{equation}
+\label{dhurdle}
+  h_2(y_i | \btheta) = 
+ \begin{cases}
+  p_i & \mbox{if } y_i=0,\\
+ (1-q_i)(1-p_i)f(y_i | m_i, r)  & \mbox{if } 1\le y_i < \mu,\\
+  q_i(1-p_i)g(y_i | \mu, \sigma, \xi)  & \mbox{if } y_i \ge \mu,
+  \end{cases}
+\end{equation}
+where $\btheta$ is the (possibly) vector-valued set of model parameters, $p_i$ is the probability that observation $i$ is governed by a zero-generating process, and $q_i$ is the probability that observation $i$ is drawn from the large-count distribution conditional on it being non-zero.
+The typical-count component is left-truncated at 0 and right-truncated at $\mu$ to maintain separability in the likelihood then normalized to ensure that each component is a density.
+
+Balderama et al 2016 considered the Generalized Pareto distribution (GPD) to describe the third component of the count data. If $y \sim \GPD(\mu, \sigma, \xi)$, then $y$ has density function
+\begin{equation}
+ g(y | \mu, \sigma, \xi) = \dfrac{\sigma^{1/\xi}}{\Paren{\sigma + \xi \times (y-\mu)}^{\frac{1}{\xi}+1}}
+\end{equation}
+for $y\ge\mu$ and $\xi \ne 0$.
+Since we are dealing with count data, we use a discretized version of the GPD distribution function, defined by taking the difference between the cumulative densities at $y+0.5$ and $y-0.5$.
+
+Notice that when $\mu=\infty$, (\ref{dhurdle}) simplifies to (\ref{nbhurdle}).
+Also notice that setting $\mu=1$ essentially removes the second component, thus all non-zero counts are modeled under the large-count distribution.
+Balderama et al 2016 defined this as the GPD-hurdle model, a new single-hurdle alternative to the negative binomial distribution for extremely right-skewed count data.
+
+%============================================
 \section[Implementing hurdlr]{Implementing \pkg{hurdlr}}
+%============================================
 
+%============================================
 \section{Discussion}
+%============================================
 
+%============================================
 \section*{References}
+%============================================
 

### 2016-10-03 paper/paper101.tex
updates
Code edits and paper additions

@@
 
-\usepackage{amssymb, amsthm, amsmath}
+\usepackage{amssymb, amsthm, amsmath, blindtext, enumitem}
 
@@
 
+%______________________________________________
+\subsection{Zero-inflated modeling}
+
+The \pkg{hurdlr} package supports zero-inflated modeling with the functions \code{zero\textunderscore nb} and \code{zero\textunderscore poisson} which, depending on usage, applies a negative binomial or Poisson distribution to count data with an excess of zeros.
+Function usage is denoted:
+
+\code{zero\textunderscore nb(y, x, size, a, b, mu.start, beta.prior.mean, beta.prior.sd, iters, burn, nthin, plots, progress.bar)}
+
+or
+
+\code{zero\textunderscore poisson(y, x, a, b, lam.start, beta.prior.mean, beta.prior.sd, iters, burn, nthin, plots, progress.bar)}
+
+with the arguments defined as follows.
+
+\begin{description}[labelindent = 1.5cm, leftmargin = 1.5cm]
+
+	\item[\code{y}] numeric response vector.
+	\item[\code{x}] numeric predictor matrix.
+	\item[\code{size}] size ($r$) parameter for NB($r$, $\mu$) likelihood distributions.
+	\item [\code{a}] shape parameter for Gamma($a$, $b$) prior distributions.
+	\item [\code{b}] rate parameter for Gamma($a$, $b$) prior distributions.
+	\item [\code{size}] size ($r$) parameter for NB($r$, $\mu$) likelihood distributions.
+	\item[\code{mu.start, lam.start}] initial value of distributional mean for NB($r$, $\mu$) or Pois($\lambda$) likelihood, respectively.
+	\item [\code{beta.prior.mean}] mean ($\mu$) for Normal($\mu$, $\sigma^{2}$) prior distributions for coefficient estimation.
+	\item [\code{beta.prior.sd}]\ standard deviation ($\sigma$) for Normal($\mu$, $\sigma^{2}$) prior distributions for coefficient estimation.
+	\item [\code{iters}] number of iterations for the Markov chain algorithm. 
+	\item [\code{burn}] numeric burn-in length.
+	\item [\code{nthin}] numeric thinning rate.
+	\item [\code{plots}] logical operator. \code{TRUE} to print plots. 
+	\item [\code{progress.bar}] logical operator. \code{TRUE} to print progress bar.
+
+\end{description}
+
+Both \code{zero\textunderscore nb} and \code{zero\textunderscore poisson} functions output a list with the elements:
+
+%______________________________________________
+\subsection{Hurdle modeling}
+
+The \pkg{hurdlr} package allows for hurdle modeling through the \code{hurdle} function.
+The \code{hurdle} function supports both single- and double-hurdle models as well as a variety of distributions for both ``typical'' distribution f(y|m, r) and the distribution for ``extreme'' values g(y|$\mu$, $\sigma$, $\xi$)
+
+{\large \textit{Forms of usage}}
+
+A \code{hurdle} function has two forms of usage; one to fit a single-hurdle model, truncated at zero, and the other to fit a double-hurdle model with an additional hurdle at some `extreme' threshold $\psi$.
+The two usages are, respectively:
+
+\code{hurdle(y, x, Inf,
+	dist, control, iters, burn, nthin, plots, progress.bar)}
+
+and
+
+\code{hurdle(y, x, hurdle,
+	dist, dist.2, control, iters, burn, nthin, plots, progress.bar)} \\
+
+with the arguments defined as follows.
+
+\begin{description}[labelindent = 1.5cm, leftmargin = 1.5cm]
+
+	\item [\code{y}] numeric response vector.
+	\item [\code{x}] optional numeric predictor matrix.
+	\item [\code{hurdle}] numeric threshold ($\psi$) for `extreme' observations of two-hurdle models. \code{Inf} for one-hurdle models.
+	\item [\code{dist}] character specification of response distribution. The function supports a choice of Poisson, negative binomial, log-normal, and generalized Pareto distributions.
+	\item [\code{dist.2}] character specification of response distribution for `extreme' observations of two-hurdle models.
+The function supports a choice of Poisson, negative binomial, log-normal, and generalized Pareto distributions.
+\code{"none"} for one-hurdle models.
+	\item [\code{control}] list of parameters for controlling the fitting process, specified by a \code{hurdle\textunderscore control()} assignment.
+	\item [\code{iters}] number of iterations for the Markov chain algorithm. 
+	\item [\code{burn}] numeric burn-in length.
+	\item [\code{nthin}] numeric thinning rate.
+	\item [\code{plots}] logical operator. \code{TRUE} to print plots. 
+	\item [\code{progress.bar}] logical operator. \code{TRUE} to print progress bar.
+
+\end{description}
+
+{\large \textit{Additional control parameters}}
+
+Many of a \code{hurdle} function's fitting parameters are defined through a \code{hurdle\textunderscore control} list.
+This includes distributional priors, hyperparameters, and MCMC tuning parameters.
+The \code{hurdle\textunderscore control} has a usage of:
+
+\code{hurdle\textunderscore control(a, b, size, beta.prior.mean, beta.prior.sd, beta.tune, pars.tune, lam.start, mu.start, sigma.start, xi.start)}
+
+with the arguments defined as follows.
+
+\begin{description}[labelindent = 1.5cm, leftmargin = 1.5cm]
+
+	\item [\code{a}] shape parameter for Gamma($a$, $b$) prior distributions.
+	\item [\code{b}] rate parameter for Gamma($a$, $b$) prior distributions.
+	\item [\code{size}] size ($r$) parameter for NB($r$, $\mu$) likelihood distributions.
+	\item [\code{beta.prior.mean}] mean ($\mu$) for Normal($\mu$, $\sigma^{2}$) prior distributions for coefficient estimation.
+	\item [\code{beta.prior.sd}]\ standard deviation ($\sigma$) for Normal($\mu$, $\sigma^{2}$) prior distributions for coefficient estimation.
+	\item [\code{beta.tune}] MCMC tuning for regression coefficient estimation.
+	\item [\code{pars.tune}] MCMC tuning for parameter estimation.
+	\item [\code{lam.start, mu.start, sigma.start, xi.start}] initial value(s) for parameter(s) of `extreme' observations distribution.
+
+\end{description}
+
+The output of the \code{hurdle} function is a list with the elements:
+
+%______________________________________________
+\subsection{An example of use}
+
+[seabird data info]
+
+The following code illustrates the use of \code{hurdle} to model the seabird data of species Sooty Shearwater.
+As indicated by Table 1, this data is both zero-inflated and exceptionally over-dispersed, and may best be fit using a GPD-hurdle model.
+The negative binomial distribution is specified for the ``typical'' observed counts, and ``extreme'' observations are modeled by the discrete generalized Pareto distribution.
+The hurdle to delineate ``extreme'' values is set as the 90th percentile of the non-zero counts (in this case, 14).
+The MCMC algorithm is run for 25,000 iterations with a 5,000 burn-in and thinning interval of 5.
+Uninformative priors and tuning parameters are used for all pieces of the model.
+
+\begin{tabular}{ |l|r| }
+	\hline
+	\multicolumn{2}{|c|}{\textbf{Avian Counts: Sooty Shearwater}} \\
+	\hline
+	\textbf{Count} &\textbf{ Frequency} \\ \hline
+	0 & 33503 \\ \hline
+	1 - 10 & 412 \\ \hline
+	11 - 100 & 88 \\ \hline
+	101 - 500 & 16 \\ \hline
+	501 + & 3 \\ \hline
+\end{tabular}
+
+Table 1: Frequency seabird counts for the species Sooty Shearwater.
+Non-zero counts have a range of 1 to 700 as well as a mean and median of 9.55 and 2, respectively.
+The resulting distribution is extremely right-skewed.
+
+How do I input/output code???
+
+\begin{CodeChunk}
+	\begin{CodeInput}
+
+#Load in seabird (species "sosh")
+load(url("http://math.luc.edu/~ebalderama/stat388bayes_resources/sb/data/sosh.RData"))
+
+#Set a value for the second hurdle
+hurd <- quantile(Y[Y!=0], prob = 0.9)
+
+#Set control parameters		
+control <- hurdle_control(a = 1, b = 1, size = 1, beta.prior.mean = 0, beta.prior.sd = 10, pars.tune = 0.2, beta.tune = 1, lam.start = 1, mu.start = 1, sigma.start = 1, xi.start = 1)
+
+#Run function for a double-hurdle model
+d.hurdle <- hurdle(y = Y, x = X[,-1], hurd = hurd, dist = "nb", dist.2 = "gpd", control = control, iters = 25000, burn = 5000, nthin = 5, plots = T, progress.bar = T)
+	\end{CodeInput}
+\end{CodeChunk}
+
 %============================================
